
---
title: "Module 7: fastlink, Part I"
author: "Rebecca C. Steorts"
output: 
     beamer_presentation:
      includes: 
          in_header: custom2.tex
font-size: 8px
---

Reading
===

- Binette and Steorts (2020)
- Edmorando et al. (2020)
- Fellegi and Sunter (1969)


# Probabilistic Entity Resolution 

While Fellegi and Sunter (1969) have provided a framework for probabilistic entity resolution, there are few implmentations that scale to large data sets. 

# Agenda

- We review fastlink, Edmorando et al. (2020)
- We illustrate a toy example on \texttt{RLdata10000}

# fastlink

- Edmorando et al. (2020) developed fastlink
a scalable implementation of the FS method. 

- In addition, the authors incorporated
auxiliary information such as population name frequency and migration rates.

- The authors used parallelization and hashing to merge millions of records in
a near real-time on a laptop computer, and provided open-source software of
their proposed methodology. 

# Agreement Patterns

\begin{itemize}
\item Assume two data sets ($A$ and $B$) with overlapping variables in common (such as name, gender, address, etc.)
\item Define an agreement value in field $a$ for record pair $(i,j):$
\end{itemize}

$$\rho_a(i,j) \ = \ \left\{\begin{array}{l}   \color{blue}{\texttt{agree}} \\ \\
    								     \color{blue}{\texttt{disagree}} \end{array}
                               \right.$$
                               
                               


# Agreement Patterns

\begin{table}[]
\begin{tabular}{ccccccc}
    &  & First & Last &  Age & Street \\
  \hline
    & \multicolumn{3}{l}{Data set $\mathcal{A}$} \\
    &  1 &  \texttt{James}  	& \texttt{Smith}  	& \texttt{35} & \bf{{\texttt{Devereux St.}}} \\
    &  \multicolumn{3}{l}{Data set $\mathcal{B}$} \\
    &  7 &  \texttt{James} 	& \texttt{Smit} 		& \texttt{43} & \bf{{\texttt{Dvereux St.}}}  \\
	 \hdashline
    & & \color{blue} {\texttt{agree}} 	& \color{blue} {\texttt{agree}}  	& \color{blue} {\texttt{disagree}} 		&   \color{blue} {\bf{{\texttt{agree}}}} \\  \hline
\end{tabular}
\end{table}

\vspace*{4em}
\pause

\begin{center}
\textbf{Agreement pattern} $\gamma(i, j) = \{\gamma_1(i,j), \gamma_2(i,j), \ldots, \gamma_K(i,j) \}$ 
\end{center}  


# Agreement Patterns

- We **observe** the agreement patterns $\gamma(i,j)$ 
- We **do not observe** the matching status 

$$C_{i,j} \ = \ \left\{\begin{array}{l}   \color{blue}{\texttt{non-match}} \\ \\
    								     \color{red}{\texttt{match}} \end{array}
                               \right.$$



# \texttt{fastLink} Model

\begin{center}
\begin{eqnarray*}
  C(i, j) & \iid & \text{Bernoulli}(\mu) \\ 
  \rho(i,j) \mid C(i, j) = \text{\color{blue}{non-match}} & \iid &  \mathcal{F}({ \color{blue}{\pi}_{\text{\color{blue}{NM}}} }) \\ 
  \rho(i,j) \mid C(i, j) = \text{\color{white}{non-}\color{red}{match}} & \iid &  \mathcal{F}( {\color{red}{\pi}_{\text{\color{red}{M}}} }),
\end{eqnarray*} 
\end{center}

where $\lambda$, ${\color{red}{\pi}_{\text{\color{red}{M}}}}$, ${\color{blue}{\pi}_{\text{\color{blue}{NM}}}}$  are estimated via the EM algorithm


# \texttt{fastLink} package

- Available on CRAN
- We investigate it first on RLdata10000
- We assume no blocking 

# fastlink in R

```{r, message=FALSE}
# load libraries 
library(fastLink)
library(RecordLinkage)
```

# RLdata10000

\footnotesize
```{r}
# load RLdata10000
records <- 
  read.table("data/RLdata10000.csv", 
             sep=",", header=TRUE)
head(records, 4)
```

# RLdata10000

```{r}
# Number of unique records 
length(unique(records$ent_id))
```

# Linkage Fields

\footnotesize
```{r}
# linkage fields 
linkageFields <- c("fname_c1", "lname_c1", "by", "bm", "bd")
```

# Exact Matching 

```{r}
# perform exact matching 
exact.match <- merge(records, records, 
                     by = linkageFields)

# number of self-matches
sum(exact.match$rec_id.x == exact.match$rec_id.y)

# number of non-self matches
sum(exact.match$rec_id.x != exact.match$rec_id.y)
```

# Who are the non-self matches?

```{r}
head(exact.match[exact.match$rec_id.x 
                 != exact.match$rec_id.y,
                 c(linkageFields)], 4)

```


# Preparation 

```{r}
# linkage fields 
linkageFields <- c("fname_c1", 
                   "lname_c1", "by", "bm", "bd")

# string distance fields
stringDistFields <- c("fname_c1", "lname_c1")

# partial distance fields (fields where we allow
# for agree, disagree, and partially agree)
partialMatchFields <- c("fname_c1", "lname_c1")
```

# Run \texttt{fastLink}

```{r, message=FALSE}
out <- fastLink(dfA = records,
                dfB = records,
                varnames = linkageFields,
                # JW by default
                stringdist.match = stringDistFields, 
                partial.match = partialMatchFields,
                # JW cutoffs
                cut.a = 0.94, cut.p = 0.84, 
                # 1-to-1 match
                dedupe = FALSE)
```

# \texttt{fastLink} Objects

\texttt{fastLink} has the following objects as output: 

\small
```{r}
names(out)
```

# Who is matched?

The indices of each matched pair can be found in out$matches

```{r}
head(cbind(out$matches$inds.a, out$matches$inds.b), 6)
```

# Counting Patterns 

- Counts and FS weights for each patterns can be found in `out$EM$patterns.w`

- Legend: 2 = Agree; 1 = Partially Agree; 0 = Disagree

\footnotesize
```{r}
tail(out$EM$patterns.w[, 1:7])
```

Matching Threshold
===

By default the **matching threshold 0.85**, but it can be easily changed:

```{r}
out <- fastLink(dfA = records, dfB = records, 
                varnames = linkageFields,
                stringdist.match = stringDistFields, 
                partial.match = partialMatchFields,
                cut.a = 0.94, cut.p = 0.84,
                # Matching threshold dedupe = FALSE
                threshold.match = 0.90,)
```

# Loading Packages

```{r}
source("evaluationMetrics.R")
#library(exchangeableER)
library(magrittr)
library(data.table)
```

# Setup

```{r}
trueMembership <- records$ent_id
recordIds <- records$rec_id
numRecords <- dim(records)[1]
matches <- 
  data.table(cbind(out$matches$inds.a,     
                   out$matches$inds.b))
head(matches)
dim(matches)[1]
```

# True Positives, False Positives, and False Negatives
```{r}
## True Positives, False Positives, and False Negatives:
TP <- sum(records$ent_id[matches$V1] 
          == records$ent_id[matches$V2])
FP <- sum(records$ent_id[matches$V1] 
          != records$ent_id[matches$V2])
FN <- dim(matches)[1] - TP
```

# FDR and FNR
```{r}
## False Discovery Rate
FDR <- round(FP/(FP + TP), 2)
FDR

## False Negative Rate
FNR <- round(FN/dim(matches)[1], 2)
FNR
```

# Precision and Recall

```{r}
precision <- 1 - FDR

recall <- 1 - FNR 

f1 <- (2.0*TP)/(2.0*TP+FP+FNR)

cbind(precision, recall, f1)
```

# Summary

- We have introduced the basics of \texttt{fastLink}
- How would one incorporate blocking to scale to larger data sets? 
- How would we apply this package to a real data set beyond \texttt{RLdata10000}?
