
---
title: "Module X: Blocking"
author: "Rebecca C. Steorts"
output: 
     beamer_presentation:
      includes: 
          in_header: custom2.tex
font-size: 8px
---



Agenda
===

- Data Cleaning Pipeline
- Blocking
- Traditional Blocking
- Probabilistic Blocking

Load R packages
===

```{r, echo=FALSE }
library(RecordLinkage)
```

Data Cleaning Pipeline
===

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{finalFigures/pipeline}
    \caption{Data cleaning pipeline.}
    \end{center}
\end{figure}

Blocking
===

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{finalFigures/block.png}
    \caption{Left: All to all record comparison. Right: Example of resulting blocking partitions. }
    \end{center}
\end{figure}


Blocking
===

- Blocking partitions similar records into partitions/blocks.

- ER is only performed within each blocks. 

Traditional Blocking
===

- A deterministic (fixed) partition is formed based upon the data. 
- A partition is created by treating certain fields that are thought to be nearly error-free as fixed.
- Benefits:  simple, easy to understand, and fast to implement. 
- Downsides: the blocks are treated as error free, which is not usually accurate and can lead to errors in the ER task that cannot be accounted for. 

Example: Blocking on date of birth year. 

Probabilistic Blocking 
===

- A probability model is used to cluster the data into blocks/partitions. 

Example: Fellegi-Sunter (1969), or Locality Sensitive Hashing  

Under both blocking approaches, record pairs that do not meet the blocking criteria are automatically classified as non-matches.

# Example: Traditional blocking



\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{FinalFigures/noblocking_plot.pdf}
    \includegraphics[width=0.45\textwidth]{FinalFigures/blocking_plot.pdf}
    \caption{All-to-all record comparisons (left) versus partitioning records into blocks by lastname initial and comparing records only within each partition (right).}
    \end{center}
\end{figure}

# Example: RLdata500

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(RecordLinkage)
data(RLdata500)
head(RLdata500)
```

# RLdata500 (Continued)

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Total number of all to all record comparisons
choose(500,2)
```

# RLdata500 (Continued)

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Block by last name initial  
last_init <- substr(RLdata500[,"lname_c1"], 1, 1)
head(last_init)

# Total number of blocks
length(unique(last_init))
```

# RLdata500 (Continued)

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Total number of records per block
recordsPerBlock <- table(last_init)
head(recordsPerBlock)

# Block sizes can vary 
summary(as.numeric(recordsPerBlock))
```

# RLdata500 (Continued)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

# Total number of records pairs per block
sapply(recordsPerBlock, choose, k=2)

# Reduction on comparison space
sum(sapply(recordsPerBlock, choose, k=2))
```

# RLdata500 (Continued)

What is the overall dimension reduction form the orginal space to the reduced space induced by blocking? 

Recall the original space of comparisons was 
```{r, eval=TRUE, message=FALSE, warning=FALSE}
choose(500,2)
```

We have reduced the number of comparisons to 
```{r, eval=TRUE, message=FALSE, warning=FALSE}
sum(sapply(recordsPerBlock, choose, k=2))
```

\textcolor{red}{Calculate the RR.}

# Pairwise Evaluation Metrics

Calculate the Pairwise Precision and Recall. 

# Case Study to El Salvador 

We return to the case study on El Salvador, where we will investigate deterministic blocking as done in Sadinle (2014). 

# Task 1

Implement the blocking procedure from Sadinle (2014), where the blocking criterion is XXX. 

# Task 2

Explain why you think the author choose this blocking criterion. 

# Task 3

What is the reduction ratio, precision, and recall assuming that the ground truth is true in this situation? 

# Task 4

Can you come up with a better blocking criterion for this data set that is deterministic? 


