
---
title: "Homework 1, STA 490/690"
author: "Rebecca C. Steorts"
output: 
     pdf_document:
      includes: 
          in_header: custom2.tex
font-size: 8px
---

***General instructions for homeworks***: Please follow the uploading file instructions according to the syllabus. Your code must be completely reproducible and must compile. 

***Advice***: Start early on the homeworks and it is advised that you not wait until the day of as these homeworks are meant to be longer and treated as case studies.

***Commenting code***
Code should be commented. See the Google style guide for questions regarding commenting or how to write 
code \url{https://google.github.io/styleguide/Rguide.xml}. No late homework's will be accepted.

***R Markdown Test***

0. Open a new R Markdown file; set the output to HTML mode and "Knit". This should produce a web page with the knitting procedure executing your code blocks. You can edit this new file to produce your homework submission.

***Working with data***

\textbf{Total points on assignment: 5 (reproducibility) + 10 points for the assignment.}

\section*{El Salvador Civil War}

Recall that between 1980 and 1991, the Republic of El Salvador witnessed a civil war between the central government, the left-wing guerrilla Farabundo Marti National Liberation Front (FMLN), and right-wing paramilitary death squads. After the peace agreement in 1992, the United Nations created a Commission on the Truth (UNTC) for El Salvador, which invited members of Salvadoran society to report war-related human rights violations, which mainly focused on killings and disappearances.In order to collect such information the UNTC invited individuals through newspapers, radio, and television advertisements to come forward and testify.  The UNTC opened offices through El Salvador where witnesses could provide their testimonials, and this resulted in a list of potential victims with names, date of death, and reported location.

In this assignment, you will explore the UNTC data set to get a better understanding of how to work with real data versus toy data. Let's read in the data. 

```{r, message=FALSE}
library(knitr)
library(RecordLinkage)
```

```{r}
# read in data
df <- read.csv("./sv-mauricio/sv-mauricio.csv")
head(df)
dim(df)
```

Next, let's filter out any records that do not have ground truth information. 

```{r}
ent_id <- df$HandID
# Filter out records with ground truth, leaving dept 1 and 7
df <- df[!is.na(ent_id),]
ent_id <- ent_id[!is.na(ent_id)]
head(df)
tail(df)
dim(df)
```

Observe that we are only considering two municipalities in El Salvador now, which is what was considered in Sadinle (2014). 


\section*{Task 1: Similarity of Hispanic Names}
\textbf{Consider the similarity of first name and last name. Consider looking at the Edit distance and comment on how this works. Below we explain a recently proposed hybrid metric for Hispanic names. Consider this using the code provided. Explain your findings.}

\paragraph{Background on the Monge Elkan Distance Metric}

It will be useful (for this case study) to utilize a hybrid distance measure for 
comparing textual strings containing multiple words (tokens). 
A hybrid distance measure accounts for differences between tokens, while 
allowing for fuzzy matching between tokens.
The measure we describe here resembles a hybrid similarity measure 
proposed by Monge (1996)  for attribute matching. As shown in Marchant et. al. (2020)
this metric attempts to match the tokens in each string while incorporating penalties for ``missing'' 
tokens. 

We describe the measure with reference to the following example.
Suppose we have two strings to compare:
$y =$ "University of California, San Diego'' and 
$x =$ "Univ. Calif., Sna Diego''. 

The two strings clearly refer to the same entity, however the latter 
string is abbreviated and has a typographical error.
We begin by splitting each string into tokens:
\begin{align*}
  \vec{y} &= [\text{``University'', ``of'', ``California,'', ``San'', ``Diego''}], \\
  \vec{x} &= [\text{``Univ.'', ``Calif.,'', ``Sna'', ``Diego''}].
\end{align*}

As explained in Marchant et al. (2020), to compute the distance between $\vec{y}$ and $\vec{x}$, one must proceed in three steps:
\begin{enumerate}
  \item If $\vec{y}$ and $\vec{x}$ do not contain the same number of 
  tokens, we append special \texttt{NA} tokens to the shorter token 
  vector to make $|\vec{y}| = |\vec{x}|$. 
  For the above example, we append an \texttt{NA} token to $\vec{x}$.
  \item Next we view $\vec{y}$ and $\vec{x}$ as two independent parts 
  of a bipartite graph.
  We solve the minimum weight matching problem for the bipartite graph, 
  where the weights between nodes (tokens) are defined by an \emph{inner} 
  distance measure $\operatorname{dist}_\mathrm{inner}(y_i, x_j)$ for 
  $y_i \in \vec{y}$ and $x_i \in \vec{x}$. 
  Note that $\distfn_\mathrm{inner}(y_i, x_j)$ must return valid 
  distances when either $y_i$ or $x_j$ is an \texttt{NA} token. 
  We represent the solution of the minimum weight bipartite matching 
  problem as an edge set 
  $M = \{(y_i \leftrightarrow x_j) : i \in 1, \ldots, |\vec{y}|\}$. 
  For the above example, we obtain  
  $M = \{(\text{``University''} \leftrightarrow \text{``Univ.''}), 
         (\text{``of''} \leftrightarrow \texttt{NA}), 
         (\text{``California,''} \leftrightarrow \text{``Calif.,''}), 
         (\text{``San''} \leftrightarrow \text{``Sna''}), 
         (\text{``Diego''} \leftrightarrow \text{``Diego''})\}$ 
  using the inner distance measure defined below.
  \item Finally we define the hybrid distance between $y$ and $x$ 
  as an average over the minimum matching weights:
  \begin{equation*}
    \distfn(y, x) = \frac{1}{|M|} \sum_{(y_i \leftrightarrow x_j) \in M} \distfn_\mathrm{inner}(y_i, x_j).
  \end{equation*}
  For the above example, we obtain $\distfn(y, x) = 0.62$\footnote{
    Using $d_\mathrm{miss,l} = \infty$, $d_\mathrm{miss,r} = 0$, 
    $d_\mathrm{abbr,l} = 0.1$, $d_\mathrm{abbr,r} = 1$.
  } which is an good result given that $y$ and $x$ have the same 
  semantic meaning.
  In comparison, the Levenshtein (edit) distance 
  $\distfn_\mathrm{Ed}(y,x) = 12$ does not reflect the semantic 
  closeness of $y$ and $x$.
\end{enumerate}

The inner distance measure $\distfn_\mathrm{in}$ plays a crucial role in 
the performance of the hybrid distance. The authors utilize a 
modified Levenshtein distance which handles \texttt{NA} tokens and detects abbreviations.

Concretely, they set
\begin{equation*}
  \distfn_\mathrm{inner}(y, x) = \begin{cases}
    d_\mathrm{miss,l}, & \text{if } y = \mathtt{NA}, \\
    d_\mathrm{miss,r}, & \text{if } x = \mathtt{NA}, \\
    d_\mathrm{abbr,l} \cdot \distfn_\mathrm{Ed}(y, x), & \text{if } y \text{ abbreviates } x, \\
    d_\mathrm{abbr,r} \cdot \distfn_\mathrm{Ed}(y, x), & \text{if } x \text{ abbreviates } y, \\
    \distfn_\mathrm{Ed}(y, x) & \text{otherwise}.
  \end{cases}
\end{equation*}
where $d_\mathrm{miss,l}$, $d_\mathrm{miss,r}$, $d_\mathrm{abbr,l}$, 
$d_\mathrm{abbr,l}$ are positive constants and $\distfn_\mathrm{Ed}$ 
is the Levenshtein distance.





```{r}
# Normalized Levenshtein similarity function used below
unitLevenshteinSimilarity <- function(v1, v2) {
  totalLength <- matrix(nchar(v1), nrow=length(v1), ncol=length(v2))
  totalLength <- sweep(totalLength, 2, nchar(v2), FUN = "+")
  dist <- adist(v1, v2)
  ifelse(totalLength > 0, 1.0 - 2.0 * dist / (totalLength + dist) , 1.0)
}
```





```{r}
#' Similarity function for Hispanic names based upon the Monge Elkan metric
#'
#' @param x a character vector
#' @param y a character vector
#' @param sep separator for tokens/words (uses white space by default)
#' @param knownTokens a character vector of known tokens (default is NULL)
#' @returns a length(x) × length(y) similarity matrix
unitHispanicSimilarity <- function(x, y, sep = '\\s+', knownTokens = NULL) {
  # Split into tokens (words)
  tokens1 <- strsplit(x, sep)
  tokens2 <- strsplit(y, sep)
  
  # Preallocate similarity matrix for output
  out <- matrix(0.0, nrow = length(tokens1), ncol = length(tokens2))
  
    if (!is.null(knownTokens)) {
    # Convert known tokens to environment for faster look-up
    knownList <- setNames(replicate(length(knownTokens), 1, simplify = FALSE), knownTokens)
    knownEnv <- list2env(knownList, hash = TRUE, size = length(knownList))
  }
  
  # Function to compute the symmetrized Monge-Elkan similarity for a single 
  # pair of tokens
  meSim <- function(t1, t2) {
    maxSim1 <- numeric(length=length(t1))
    knownDistinct1 <- logical(length=length(t1))
    maxSim2 <- numeric(length=length(t2))
    knownDistinct2 <- logical(length=length(t2))
    for (i in seq_along(t1)) {
      for (j in seq_along(t2)) {
        sim <- unitLevenshteinSimilarity(t1[i], t2[j])
        bothKnownDistinct <- FALSE
        if (!is.null(knownTokens) && t1[i] != t2[j] &&
            exists(t1[i], envir = knownEnv, inherits = FALSE) && 
            exists(t2[i], envir = knownEnv, inherits = FALSE)) {
          bothKnownDistinct <- TRUE
        }
        if (sim > maxSim1[i]) { maxSim1[i] <- sim; knownDistinct1[i] <- bothKnownDistinct }
        if (sim > maxSim2[j]) { maxSim2[j] <- sim; knownDistinct2[j] <- bothKnownDistinct }
      }
    }
    maxSim1 <- ifelse(knownDistinct1, 0, maxSim1)
    maxSim2 <- ifelse(knownDistinct2, 0, maxSim2)
    # Symmetrize
    return(max(length(t1)/sum(1.0/maxSim1), length(t2)/sum(1.0/maxSim2)))
  }
  
  # Function to compute an asymmetric similarity for a single pair of tokens 
  asymSim <- function(t1, t2) {
    if (length(t1) < length(t2)) {
      # If t2 contains extra tokens, similarity is zero (can't distort 
      # true name by adding names)
      return(0)
    } else {
      # Get symmetrized Monge-Elkan similarity
      me <- meSim(t1, t2)
      # Assign 0.95 weight to Monge-Elkan and 0.05 weight to num. tokens 
      # similarity
      #return(1.0/(0.95/me + 0.05*length(t1)/length(t2)))
      return(me)
    }
  }
  
  # Loop over all combinations in input character vectors
  for (i in seq_len(length(tokens1))) {
    for (j in seq_len(length(tokens2))) {
      out[i, j] <- asymSim(tokens1[[i]], tokens2[[j]])
    }
  }
  
  return(out)
}
```

\section*{Solution to Task 1}

Consider comparing the following two names using Edit distance: `r df$firstname[1]` and `r df$firstname[2]`. We find that using the following code below that the distance is just `r levenshteinSim(df$firstname[1], df$firstname[2])`. This seems quite reasonable given that both names are quite different.

```{r, message=FALSE}
levenshteinSim(df$firstname[1], df$firstname[2])
```

Consider comparing the following two names using Edit distance: `r df$firstname[731]` and `r df$firstname[732]`.
We find the Edit distance is `r levenshteinSim(df$firstname[731], df$firstname[732])`.

We have now seen an interesting case. Hopefully, you have noticed that Hispanic first names, have two tokens, and thus, are much longer than Western names. Perhaps we might want to change the distance function. 

Now when comparing `r df$firstname[1]` and `r df$firstname[2]` under the Monge Elkan metric our score is
`r unitHispanicSimilarity(df$firstname[1], df$firstname[2])`. 

Now when comparing `r df$firstname[731]` and `r df$firstname[732]` under the Monge Elkan metric our score is
`r unitHispanicSimilarity(df$firstname[731], df$firstname[732])`. 



\section*{Task 2}

\textbf{How does exact matching work on this data set? What about off by one matching? Be sure to provide the precision and recall. Hint: Be sure to work on the modified data set below as I have removed columns that would not be wise for comparing, such as the record label.}

```{r}
head(df)
head(df_new <- df[,3:8,10])
```

\section*{Solution Task 2}

\section*{Task 3}

\textbf{How would you build a decision rule for matches/non-matches based upon scoring rules. What would your scoring rule be? Write this up as an algorithm.}

\section*{Solution Task 3}


\section*{Task 4}

\textbf{Code up your algorithm in Task 3 and provide the precision and recall. Did your method do better or worse than exact matching?}

\section*{Solution Task 4}

\section*{Task 5}

\textbf{Give insights into how you might be able to improve deterministic approaches moving forward if you re-did your analysis. What advice would you give to a new member that is just joining the project after working on this project (assume that they have just joined your team and your job is to bring them up to speed).}

\section*{Solution Task 5}


