
---
title: "Module 8: Bayesian Fellegi and Sunter"
author: "Rebecca C. Steorts"
output: 
     beamer_presentation:
      includes: 
          in_header: custom2.tex
font-size: 8px
---


# BDD package 

- Marchant, Rubinstein, and Steorts (2021) have provided a package for implementing Sadinle (2014), which is referred to as BDD (Bayesian Duplicated Detection). 

# RLdata500 data set

- We investigate it on the RLdata500 data set. 

\footnotesize
```{r, message=FALSE}
library(magrittr)    # pipe operator (must be loaded before BDD)
library(comparator)  # normalized Levenshtein distance
library(clevr)       # evaluation functions
library(BDD)
RLdata500[['rec_id']] <- seq.int(length.out=nrow(RLdata500))
head(RLdata500)
```

# Distance functions

```{r, eval=TRUE, message=FALSE, warning=FALSE}
scoring_fns <- list(
  fname_c1 = Levenshtein(normalize = TRUE),
  lname_c1 = Levenshtein(normalize = TRUE),
  by = function(x, y) abs(x - y),
  bm = function(x, y) abs(x - y),
  bd = function(x, y) abs(x - y)
)
```

# Breaks

For each scoring function above, we provide a breaks vectors, which specifies the discrete levels of agreement (from 'high' agreement to 'low').

```{r, eval=TRUE, message=FALSE, warning=FALSE}
scoring_breaks <- list(
  fname_c1 = c(-Inf,.05,.15,.3,Inf),
  lname_c1 = c(-Inf,.05,.15,.3,Inf),
  by = c(-Inf,0,1,3,Inf),
  bm = c(-Inf,0,1,3,Inf),
  bd = c(-Inf,0,2,7,Inf)
)
```

# Comparison vectors 

- Now we are ready to compute the attribute comparison scores for the record pairs. 

- Since this is a small data set, we consider all pairs using the  `pairs_all` function. 

- For larger data sets, blocking/indexing is recommended using `pairs_hamming`, ` pairs_fuzzyblock`, or a custom blocking function. 

# Comparison vectors 


```{r, eval=TRUE, message=FALSE, warning=FALSE}
pairs <- pairs_all(RLdata500$rec_id) %>%
  compute_scores(RLdata500, scoring_fns, id_col = 'rec_id') %>% 
  discretize_scores(scoring_breaks)
```

# Speeding up inference 

- To speed up inference, we only consider a subset of the pairs as  candidate matches. 

- Specifically, we consider pairs that have a strong 
agreement on name (accounting for missing names).

# Speeding up inference 

To speed up inference, we only consider a subset of the pairs as candidate matches. 

Specifically, we consider pairs that have a strong agreement on name (accounting for missing names).

\vspace*{1em}

\tiny
```{r, eval=TRUE, message=FALSE, warning=FALSE}
pairs[['candidate']] <- (pairs$fname_c1 < 4) & (pairs$lname_c1 < 4) | 
                            is.na(pairs$fname_c1) | is.na(pairs$lname_c1) 
```

# priors on $m$ and $u$ probabilities

- Next we specify the priors on the m* and u* probabilities for each 
attribute and agreement level. 

- `lambda` contains the lower truncation 
points for the truncated Beta priors on the m* probabilities. `alpha1` and `beta1` are the shape parameters for the truncated Beta distributions to the `BDD` function below.

- The Beta priors on the u* probabilities are uniform by default but can be adjusted using the alpha0 and beta0 arguments.

# priors on $m$ and $u$ probabilities

```{r, eval=TRUE, message=FALSE, warning=FALSE}
lambda <- list(
  fname_c1 = c(0.8,0.85,0.99),
  lname_c1 = c(0.8,0.85,0.99),
  by = c(0.8,0.85,0.99),
  bm = c(0.8,0.85,0.99),
  bd = c(0.8,0.85,0.99)
)

```

# Intialization  and Gibbs sampler 

Finally we initialize the model and run inference using Markov chain  Monte Carlo.

```{r, eval=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
model <- BDD(pairs, lambda, 
             id_cols = c("rec_id.x", "rec_id.y"), 
             candidate_col = "candidate")
fit <- run_inference(model, 100, 
                     thin_interval = 10, 
                     burnin_interval = 100)
# Completed sampling in 0.7 seconds
```

# Posterior samples

- The posterior samples of the linkage structure can be accessed by calling 
`extract(fit, "links")`. 

- However, these samples only cover the records that were considered as 
candidate pairs. 

- We can obtain samples of the complete linkage structure (for all records) 
using the following function.

```{r, eval=TRUE}
links_samples <- 
  complete_links_samples(fit, RLdata500$rec_id)
```

# Traceplot of the cluster sizes

```{r, eval=TRUE, echo=FALSE}
n_clusters <- extract(fit, "n_clusters")
m_probs <- extract(fit, "m")
u_probs <- extract(fit, "u")
plot(n_clusters)
```

# Traceplot of the m probabilities 

```{r, echo=FALSE}
plot(m_probs)
```

# Traceplot of the u probabilities

```{r, echo=FALSE}
plot(u_probs)
```

# Pairwise Evaluation Metrics (Boxplots)

```{r, eval=TRUE, echo=FALSE}
n_records <- nrow(RLdata500)
true_pairs <- membership_to_pairs(identity.RLdata500)
metrics <- apply(links_samples, 1, function(links) {
  pred_pairs <- membership_to_pairs(links)
  pr <- precision_pairs(true_pairs, pred_pairs)
  re <- recall_pairs(true_pairs, pred_pairs)
  c(Precision = pr, Recall = re)
}) %>% t()
boxplot(metrics)
```

# Your Turn and Discussions

- What other evaluation metrics would you look at? 

- Work on these either with a partner or in your free time? 

- What types of evaluations can we look at given the fact that this is an unsupervised problem? 

# Your Turn and Discussions

- You might think about how you would try and replicate the analysis that Sadinle did in his 2014 paper given this package. 
*
- Can you think of a simulation study to study the sensitivity of the method? 

- If you like this method, go read Sadinle (2017) and Sadinle (2018)! 
