
---
title: "Homework 4: Entity Resolution Approaches Applied to the El Salvodoran Conflict "
author: "Rebecca C. Steorts, STA 490/690"
output: 
     pdf_document:
      includes: 
          in_header: custom2.tex
font-size: 8px
---

\textbf{Total points on assignment: 5 (reproducibility) + 10 points for the assignment.}

\section*{El Salvador Civil War}

We will continue with our exploration of the UNTC data set from El Salvador. 

```{r, message=FALSE}
library(knitr)
library(RecordLinkage)
```

```{r}
# read in data
df <- read.csv("../sv-mauricio/sv-mauricio.csv")
ent_id <- df$HandID
# Filter out records with ground truth, leaving dept 1 and 7
df <- df[!is.na(ent_id),]
ent_id <- ent_id[!is.na(ent_id)]
new_df <- df[,c(3:8,10)]
head(new_df)
```


Recall that we are only considering two municipalities in El Salvador now, which is what was considered in Sadinle (2014). 

\section*{Entity resolution applied to the El Salvadoran conflict}

\textbf{In this assignment, you should explore entity resolution approaches and how they work on the El Salvadoran conflict. Specifically, each team should investigate a different entity resolution method (citing the methodology or paper) and explain their results/evaluations.}



