
---
title: "Module 6: Fellegi-Sunter Method Applied to RLdata500"
author: "Rebecca C. Steorts"
institute: joint with Olivier Binette
output: 
     beamer_presentation:
      includes: 
          in_header: custom2.tex
font-size: 8px
---

## Agenda 




## Load packages

```{r, echo=FALSE}
library(RecordLinkage)
```

## RLdata500

```{r}
data(RLdata500)
head(RLdata500)
head(identity.RLdata500)
```

## Comparison Vectors

Why do we build record pairs of comparison vectors? 

\textbf{Answer}: Reduce the total number of record comparisons.

## Comparison Vectors

How do we build record pairs of comparison vectors?

\vspace*{1em}

\textbf{Answer}: Use the `compare.dedup` function. 

## Comparison Vectors


```{r}
# create comparison vectors
rpairs <- compare.dedup(RLdata500, 
                        identity = identity.RLdata500)
```

## Comparison Vectors 

\footnotesize
```{r}
# inspect comparison vectors
rpairs$pairs[1:5,]
```

## Blocking

Blocking is the reduction of the amount of data pairs through focusing on specified agreement patterns.

Blocking is a common strategy to reduce computation time and memory consumption by only comparing records with equal values for a subset of attributes, called blocking fields. 

## Blocking

A blocking specification can be supplied to the `compare` function via the argument `blockfld`. 

We will consider a blocking pattern where two records must agree in either the **first component of the first name** or **full date of birth**.

## Blocking and Comparison Vectors

```{r}
# blocking and comparison vectors
rpairs <- compare.dedup(RLdata500, 
                        blockfld = list(1,5:7),
                        identity = identity.RLdata500)
```

## Blocking and Comparison Vectors

\footnotesize
```{r}
# inspect comparison vectors
rpairs$pairs[c(1:3, 1203:1204),]
```


Observe that these records agree on first name but not date of birth (as designed).

# String Comparators




